---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sparsediscrim

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![Codecov test coverage](https://codecov.io/gh/topepo/sparsediscrim/branch/main/graph/badge.svg)](https://codecov.io/gh/topepo/sparsediscrim?branch=main)
[![R-CMD-check](https://github.com/topepo/sparsediscrim/workflows/R-CMD-check/badge.svg)](https://github.com/topepo/sparsediscrim/actions)
<!-- badges: end -->


The R package `sparsediscrim` provides a collection of sparse and regularized discriminant analysis classifiers that are especially useful for when applied to small-sample, high-dimensional data sets. The package was archived in 2018 and was re-released in 2021. 

## Installation

You can install the stable version on [CRAN](https://cran.r-project.org/package=sparsediscrim):

```r
install.packages('sparsediscrim', dependencies = TRUE)
```

If you prefer to download the latest version, instead type:

```r
library(devtools)
install_github('topepo/sparsediscrim')
```

## Usage

The formula and non-formula interfaces can be used: 

```{r}
library(sparsediscrim)

data(parabolic, package = "modeldata")

qda_mod <- qda_shrink_mean(class ~ ., data = parabolic)
# or
qda_mod <- qda_shrink_mean(x = parabolic[, 1:2], y = parabolic$class)

qda_mod
```


## Classifiers

The `sparsediscrim` package features the following classifier (the R function is included within parentheses):

* [High-Dimensional Regularized Discriminant Analysis](https://arxiv.org/abs/1602.01182) (`rda_high_dim`) from Ramey et al. (2015)

The `sparsediscrim` package also includes a variety of additional classifiers intended for small-sample, high-dimensional data sets. These include:

| Classifier                                                    | Author                                                                                             | R Function |
|---------------------------------------------------------------|----------------------------------------------------------------------------------------------------|------------|
| Diagonal Linear Discriminant Analysis                         | [Dudoit et al. (2002)](http://www.tandfonline.com/doi/abs/10.1198/016214502753479248)              | `lda_diag()`     |
| Diagonal Quadratic Discriminant Analysis                      | [Dudoit et al. (2002)](http://www.tandfonline.com/doi/abs/10.1198/016214502753479248)              | `qda_diag()`     |
| Shrinkage-based Diagonal Linear Discriminant Analysis         | [Pang et al. (2009)](http://onlinelibrary.wiley.com/doi/10.1111/j.1541-0420.2009.01200.x/abstract) | `lda_shrink_cov()`    |
| Shrinkage-based Diagonal Quadratic Discriminant Analysis      | [Pang et al. (2009)](http://onlinelibrary.wiley.com/doi/10.1111/j.1541-0420.2009.01200.x/abstract) | `qda_shrink_cov()`    |
| Shrinkage-mean-based Diagonal Linear Discriminant Analysis    | [Tong et al. (2012)](http://bioinformatics.oxfordjournals.org/content/28/4/531.long)               | `lda_shrink_mean()`   |
| Shrinkage-mean-based Diagonal Quadratic Discriminant Analysis | [Tong et al. (2012)](http://bioinformatics.oxfordjournals.org/content/28/4/531.long)               | `qda_shrink_mean()`   |
| Minimum Distance Empirical Bayesian Estimator (MDEB)          | [Srivistava and Kubokawa (2007)](http://www.utstat.utoronto.ca/~srivasta/exp1.pdf)                 | `lda_emp_bayes()`     |
| Minimum Distance Rule using Modified Empirical Bayes (MDMEB)  | [Srivistava and Kubokawa (2007)](http://www.utstat.utoronto.ca/~srivasta/exp1.pdf)                 | `lda_emp_bayes_eigen()`    |
| Minimum Distance Rule using Moore-Penrose Inverse (MDMP)      | [Srivistava and Kubokawa (2007)](http://www.utstat.utoronto.ca/~srivasta/exp1.pdf)                 | `lda_eigen()`     |

We also include modifications to Linear Discriminant Analysis (LDA) with
regularized covariance-matrix estimators:

* Moore-Penrose Pseudo-Inverse (`lda_pseudo()`)
* Schafer-Strimmer estimator (`lda_schafer()`)
* Thomaz-Kitani-Gillies estimator (`lda_thomaz()`)
